S1: Extended Derivations for `A Data Driven Framework for Patient-Specific Neural Field Modelling'

### 2010-06-26 ltx/PLoS/LaTeX_Manuscript.tex
further editted discussion, around the MVAR part

The frequency analysis and model selection tools presented in Section~\ref{SpectralAnalysisSection} specify the sensor arrangement that is required for capturing the dominant cortical dynamics. Importantly, these new tools can also be used to design electrophysiological recording systems to avoid spatial aliasing. In addition, the frequency analysis highlights the relationship between the spatial width of the sensors and the \dean{noise} in the data. This analysis is becoming increasingly relevant as recording systems are becoming more sophisticated with higher density recordings~\cite{Brinkmann2009},

This is the first paper to derive a continuum neural field model-based estimation framework for intracortical connectivity from electrophysiological data. The importance of the estimation framework can be illustrated by considering that the connectivity structure at this scale can not be estimated using diffusion tensor imaging due to the near isotropic diffusion of water in grey matter~\cite{Assaf2008}.

The closest competitor to our `grey-box' integro-difference equation (IDE) based approach is perhaps the `black-box' multivariate auto-regressive (MVAR) model. Examples of MVAR models for estimating functional connectivity include Granger causality~\cite{Hesse2003}, the direct transfer function~\cite{Kaminski1991} and partial directed coherence~\cite{Sameshima1999}. A distinct advantage of the IDE-based technique proposed in this paper, is that the estimates of the connectivity structure are made independently of the number of sensors (assuming the field is sufficiently sampled). In addition, the number of parameters required to model the system is considerably less than a typical MVAR model. Furthermore, by using the neural field equations as a parametric form of the model, the parameters have physical meaning and therefore, can be clinically relevant. An advantage of the MVAR-based methods is that they allow for anisotropic connectivity, which is present in the longer range cortico-cortical fibres. A recent theoretical study~\cite{Jirsa2009} has demonstrated the importance of a multi-scale connectivity structure in generating the characteristic rhythms of the brain. A possible method of creating more precise models with this multi-scale structure maybe to estimate anisotropic cortico-cortical connectivity using a MVAR approach and intracortical connectivity using the methods proposed in this paper.

@@

Generating patient specific neural field models has the potential to have significant impact on several areas of neuroscientific research and clinical neurology. Specifically, the development of a patient-specific state-space model will allow for the application of techniques from engineering to be applied to neuro-dynamics. For example, state tracking could be used in brain-computer interface or epileptic seizure prediction applications, where specific regions of the state-space may indicate intent of movement or imminent seizures. In addition, a patient specific state-space model will allow application of control engineering techniques to robustly prevent the neuro-dynamics entering pathological regions of state-space using electrical stimulation.

>>>>

The frequency analysis and model selection tools presented in Section~\ref{SpectralAnalysisSection} specify the sensor arrangement that is required for capturing the dominant cortical dynamics. Importantly, these new tools can also be used to design electrophysiological recording systems to avoid spatial aliasing. In addition, the frequency analysis highlights the relationship between the spatial width of the sensors and the information in the data. This analysis is becoming increasingly relevant as recording systems are becoming more sophisticated with higher density recordings~\cite{Brinkmann2009},

This is the first paper to derive a continuum neural field model-based estimation framework for intracortical connectivity from electrophysiological data. The importance of the estimation framework can be illustrated by considering that the connectivity structure at this scale can not be estimated using diffusion tensor imaging due to the near isotropic diffusion of water in grey matter~\cite{Assaf2008}. The closest competitor to our `grey-box' integro-difference equation (IDE) based approach is perhaps the multivariate auto-regressive (MVAR) model, which is a `black-box' approach. Examples of MVAR models for estimating functional connectivity include Granger causality~\cite{Hesse2003}, the direct transfer function~\cite{Kaminski1991} and partial directed coherence~\cite{Sameshima1999}. A distinct advantage of the IDE-based technique proposed in this current study is that the number of parameters required to model the system is considerably less than a typical MVAR model. Furthermore, by using the neural field equations as a parametric form of a model, the parameters have physical meaning and, therefore, can be clinically relevant. An advantage of the MVAR-based methods is that they allow for anisotropic connectivity, which is present in the longer range cortico-cortical fibres. 
% A recent theoretical study~\cite{Jirsa2009} has demonstrated the importance of a multi-scale connectivity structure in generating the characteristic rhythms of the brain. A possible method of creating more precise models with this multi-scale structure maybe to estimate anisotropic cortico-cortical connectivity using a MVAR approach and intracortical connectivity using the methods proposed in this paper.

@@

\dean{There is definitely no repeat now. Also, I have commented out the speculative part about how to include anisotropy.}

Generating patient-specific neural field models has the potential to have significant impact on several areas of neuroscientific research and clinical neurology. Specifically, the development of a patient-specific state-space model will allow for the application of techniques from engineering to be applied to neuro-dynamics. For example, state tracking could be used in brain-computer interface or epileptic seizure prediction applications, where specific regions of the state-space may indicate intent of movement or imminent seizures. In addition, a patient specific state-space model will allow application of control engineering techniques to robustly prevent the neuro-dynamics entering pathological regions of state-space using electrical stimulation.
